SparseIDS: Learning Packet Sampling with Reinforcement Learning

### 2020-04-17 paper_rl/paper.tex
Small text fixes

To make \ours{} as lightweight as possible, our implementation allows to let the sampling and the classifier to be implemented in the same neural network, meaning that no overhead is introduced. With this option disabled, the computational power required is $2\times$ what would be required without the sampling. Thus, for this option, less than $50\%$ of packets must be sampled to gain a computational efficiency advantage. Fortunately, we can show that with shared weights the same accuracy can be achieved while having no computational overhead.

@@

%The results of the sampling part are discussed in this section.
We study the effect of the sparsity-accuracy tradeoff first, then we evaluate both discrete and continuous actions and, finally, we compare our results with other sampling techniques. Overall, sampling based on \gls{rl} reveals that a high number of packets can be skipped while maintaining good accuracy. Then, we inspect how the \gls{rl} chooses packets and show that certain patterns can be spotted.

@@

\begin{table}[h]
% XXX: Max: Without the vspace, the table violates the margins. WTF!!!
\vspace{0.075in}
\caption{Comparing classification metrics per flow between \gls{rl}-based sampling with continuous actions (tradeoff 0.1) and alternative strategies.}
\label{tab:results_others}
\centering
\begin{tabular*}{\columnwidth}{>{\rowmac}l @{\extracolsep{\fill}} >{\rowmac}c>{\rowmac}c>{\rowmac}c>{\rowmac}c>{\rowmac}c<{\clearrow}} \toprule
& \gls{rl} & random & every $i$th & relative first \emph{m} & first \emph{m} \\	\midrule
Sparsity & 76.3\% & 76.3\% & 76.3\% & 76.3\% & 76.3\% \\ \midrule
Accuracy & \textbf{99.4\%} & 96.9\% & 97.8\% & 97.3\% & 98.3\% \\
Precision & \textbf{98.5\%} & 92.4\% & 95.8\% & 93.8\% & 95.6\% \\
Recall & \textbf{99.0\%} & 95.6\% & 95.4\% & 95.5\% & 97.9\% \\
F1 & \textbf{99.7\%} & 94.0\% & 95.6\% & 94.7\% & 96.7\% \\
Youden & \textbf{98.5\%} & 93.0\% & 93.9\% & 93.4\% & 96.4\% \\
\bottomrule
\end{tabular*}
%}
\end{table}
%\begin{table}
%\caption{Comparing classification metrics per flow between \gls{rl}-based sampling with continuous actions (tradeoff $0.1$) and alternative strategies.} \label{tab:results_others}
%\centering
%\begin{tabular}{>{\rowmac}l >{\rowmac}c>{\rowmac}c>{\rowmac}c>{\rowmac}c>{\rowmac}c<{\clearrow}} \toprule
%& \gls{rl} & random & every $i$th & relative first \emph{m} & first \emph{m} \\	\midrule
%\setrow{\bfseries}Sparsity & 76.3\% & 76.3\% & 76.3\% & 76.3\% & 76.3\% \\ \midrule
%\setrow{\bfseries}Accuracy & 99.4\% & 96.9\% & 97.8\% & 97.3\% & 98.3\% \\
%Precision & 98.5\% & 92.4\% & 95.8\% & 93.8\% & 95.6\% \\
%Recall & 99.0\% & 95.6\% & 95.4\% & 95.5\% & 97.9\% \\
%F1 & 99.7\% & 94.0\% & 95.6\% & 94.7\% & 96.7\% \\
%Youden & 98.5\% & 93.0\% & 93.9\% & 93.4\% & 96.4\% \\
%\bottomrule
%\end{tabular}
%\end{table}

@@

\subsection{Shared weights}
\begin{table}[h]
%% XXX: Max: Without the vspace, the table violates the margins. WTF!!!
%\vspace{0.075in}
%\parbox{.45\linewidth}{
\caption{Using shared weights for all components of \ours{} resulting in no overhead in deployment compared to a classifier without our \gls{rl} framework.} \label{tab:results_shared}
\centering
\begin{tabular*}{\columnwidth}{>{\rowmac}l @{\extracolsep{\fill}} >{\rowmac}c>{\rowmac}c>{\rowmac}c>{\rowmac}c<{\clearrow}} \toprule
Tradeoff & 0.0 & 0.1 & 0.5 & 1.0 \\	\midrule
Sparsity & 0\% & 77.6\% & 77.6\% & \textbf{78.2\%} \\ \midrule
Accuracy & \textbf{99.5\%} & 99.4\% & 99.3\% & 99.0\% \\
Precision & \textbf{99.4\%} & 99.2\% & 98.6\% & 98.9\% \\
Recall & \textbf{98.5\%} & 98.2\% & 98.7\% & 96.9\% \\
F1 & \textbf{99.0\%} & 98.7\% & 98.6\% & 97.9\% \\
Youden & \textbf{98.4\%} & 97.9\% & 98.2\% & 96.6\% \\
\bottomrule
\end{tabular*}
%}
\end{table}
\autoref{tab:results_shared} shows that when sharing the neural network weights of the classifier, the actor and the critic, \ours{} achieves the same very high accuracy as in the case in which all every component has its own weights (\autoref{tab:results_tradeoff}). The achieved sparsity is even higher when using separate weights. One explanation for this might be that sharing weights has a regularizing effect. With shared weights, the \gls{rl} mechanism adds no computational complexity since all the operations in the neural network are the same for the actor and the classifier (the critic is not needed during deployment). The only part that is not shared is the last linear layer, which connects the last LSTM layer with the outputs.

@@

We show that using a shared neural network for both the \gls{ids} as well as the sampling can achieve the same classification accuracy and sparsity while not resulting in any computational overhead during deployment.

>>>>

To make \ours{} as lightweight as possible, our implementation allows to let the sampling and the classifier be implemented in the same neural network, meaning that no overhead is introduced. With this option disabled, the computational power required is $2\times$ of what would be required without the sampling. Thus, for this option, less than $50\%$ of packets must be sampled to gain a computational efficiency advantage. Fortunately, we can show that with shared weights the same accuracy can be achieved while having no computational overhead.

@@

%The results of the sampling part are discussed in this section.
%We study the effect of the sparsity-accuracy tradeoff first, then we evaluate both discrete and continuous actions, finally, we compare our results with other sampling techniques. Overall, sampling based on \gls{rl} reveals that a high number of packets can be skipped while maintaining good accuracy. Then, we inspect how the \gls{rl} chooses packets and show that certain patterns can be spotted.

@@

\begin{table}[h]
% XXX: Max: Without the vspace, the table violates the margins. WTF!!!
%\vspace{0.075in}
\caption{Comparing classification metrics per flow between \gls{rl}-based sampling with continuous actions (tradeoff 0.1) and alternative strategies.}
\label{tab:results_others}
\centering
\begin{tabular*}{\columnwidth}{>{\rowmac}l @{\extracolsep{\fill}} >{\rowmac}c>{\rowmac}c>{\rowmac}c>{\rowmac}c>{\rowmac}c<{\clearrow}} \toprule
& \gls{rl} & random & every $i$th & relative first \emph{m} & first \emph{m} \\	\midrule
Sparsity & 76.3\% & 76.3\% & 76.3\% & 76.3\% & 76.3\% \\ \midrule
Accuracy & \textbf{99.4\%} & 96.9\% & 97.8\% & 97.3\% & 98.3\% \\
Precision & \textbf{98.5\%} & 92.4\% & 95.8\% & 93.8\% & 95.6\% \\
Recall & \textbf{99.0\%} & 95.6\% & 95.4\% & 95.5\% & 97.9\% \\
F1 & \textbf{99.7\%} & 94.0\% & 95.6\% & 94.7\% & 96.7\% \\
Youden & \textbf{98.5\%} & 93.0\% & 93.9\% & 93.4\% & 96.4\% \\
\bottomrule
\end{tabular*}
%}
\end{table}
%\begin{table}
%\caption{Comparing classification metrics per flow between \gls{rl}-based sampling with continuous actions (tradeoff $0.1$) and alternative strategies.} \label{tab:results_others}
%\centering
%\begin{tabular}{>{\rowmac}l >{\rowmac}c>{\rowmac}c>{\rowmac}c>{\rowmac}c>{\rowmac}c<{\clearrow}} \toprule
%& \gls{rl} & random & every $i$th & relative first \emph{m} & first \emph{m} \\	\midrule
%\setrow{\bfseries}Sparsity & 76.3\% & 76.3\% & 76.3\% & 76.3\% & 76.3\% \\ \midrule
%\setrow{\bfseries}Accuracy & 99.4\% & 96.9\% & 97.8\% & 97.3\% & 98.3\% \\
%Precision & 98.5\% & 92.4\% & 95.8\% & 93.8\% & 95.6\% \\
%Recall & 99.0\% & 95.6\% & 95.4\% & 95.5\% & 97.9\% \\
%F1 & 99.7\% & 94.0\% & 95.6\% & 94.7\% & 96.7\% \\
%Youden & 98.5\% & 93.0\% & 93.9\% & 93.4\% & 96.4\% \\
%\bottomrule
%\end{tabular}
%\end{table}

@@

\subsection{Shared weights}
\begin{table}[h]
%% XXX: Max: Without the vspace, the table violates the margins. WTF!!!
\vspace{0.075in}
%\parbox{.45\linewidth}{
\caption{Using shared weights for all components of \ours{} resulting in no overhead in deployment compared to a classifier without our \gls{rl} framework.} \label{tab:results_shared}
\centering
\begin{tabular*}{\columnwidth}{>{\rowmac}l @{\extracolsep{\fill}} >{\rowmac}c>{\rowmac}c>{\rowmac}c>{\rowmac}c<{\clearrow}} \toprule
Tradeoff & 0.0 & 0.1 & 0.5 & 1.0 \\	\midrule
Sparsity & 0\% & 77.6\% & 77.6\% & \textbf{78.2\%} \\ \midrule
Accuracy & \textbf{99.5\%} & 99.4\% & 99.3\% & 99.0\% \\
Precision & \textbf{99.4\%} & 99.2\% & 98.6\% & 98.9\% \\
Recall & \textbf{98.5\%} & 98.2\% & 98.7\% & 96.9\% \\
F1 & \textbf{99.0\%} & 98.7\% & 98.6\% & 97.9\% \\
Youden & \textbf{98.4\%} & 97.9\% & 98.2\% & 96.6\% \\
\bottomrule
\end{tabular*}
%}
\end{table}
\autoref{tab:results_shared} shows that when sharing the neural network weights of the classifier, the actor and the critic, \ours{} achieves the same very high accuracy as in the case in which all every component has its own weights (\autoref{tab:results_tradeoff}). The achieved sparsity is even higher when using separate weights. One explanation for this might be that sharing weights has a regularizing effect.

With shared weights, the \gls{rl} mechanism adds no computational complexity since all the operations in the neural network are the same for the actor and the classifier (the critic is not needed during deployment). The only part that is not shared is the last linear layer, which connects the last LSTM layer with the outputs.

@@

We showed that using a shared neural network for both the \gls{ids} as well as the sampling can achieve the same classification accuracy and sparsity while not resulting in any computational overhead during deployment.
